---
title: "16S Microbial Analysis from lake sediments, peat, and soils in Wyoming mountains"
author: "Created by Jordan Von Eggers, modified by Ioana Stefanscu"
date: "`r Sys.Date()`"
output: pdf_document
---


## June 27, 2024 - Jordan Updating NMDS
## July 9-10, 2024 - Ioana Updating NMDS
## July 12, 2024 - Ioana Updating NMDS
## July 16, 2024 - Ioana working on shared and not shared taxa, top taxa by environment, and statistically significant taxa between environemnts
## August 2nd- Ioana updating with significant changes requested by Bryan. Get rid of downcore samples, split samples by surface and deep with a total of 5 categories,order samples in panels by latitude, rerun all analyses at OTU level first



# Load packages and report session info
```{r}
require(phyloseq)
require(tidyverse)
require(vegan)

# newly added packages that may need installing 
require(reshape2)
require(metagMisc)
require(RColorBrewer)
require(geosphere)
library(phylosmith)

# #installation for metagMisc
# install.packages("remotes")
# remotes::install_github("vmikk/metagMisc")
#remotes::install_github('schuyler-smith/phylosmith')
sessionInfo()
```


# Custom theme
```{r}
  custom_theme <- function() {
  theme_bw() +
    theme(
      text = element_text(color = "black", size = 13),
      axis.text = element_text(color = "black",size = 13),
      axis.title = element_text(color = "black",size = 13),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      legend.text = element_text(size = 13),
      legend.title = element_text(size = 13),
      strip.text = element_text(size = 13, color = "black"))}
```


# Load data
(this takes a few minutes)
```{r}
source("../3_CreatePhyloseq/3_CreatePhyloseq_ZOTUexact.R")
```


# Remove downcore samples
to only leep surface and deep samples only (get rid of downcore samples)
```{r}
ps
#phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 151945 taxa and 609 samples ]
#sample_data() Sample Data:       [ 609 samples by 24 sample variables ]
#tax_table()   Taxonomy Table:    [ 151945 taxa by 7 taxonomic ranks ]

ps<- subset_samples(ps, surface_or_deep %in% c("surface", "deep"))
ps
#phyloseq-class experiment-level object
#otu_table()   OTU Table:         [ 151945 taxa and 540 samples ]
#sample_data() Sample Data:       [ 540 samples by 24 sample variables ]
#tax_table()   Taxonomy Table:    [ 151945 taxa by 7 taxonomic ranks ]

n.samp.removed=609-540
n.samp.removed
# [1] 69

# Remove unwanted samples from metadata
metadata<- metadata%>% filter(surface_or_deep  %in% c("surface", "deep"))
``` 


#Transform and order samples
```{r}
#install.packages( "microViz", repos = c(davidbarnett = "https://david-barnett.r-universe.dev", getOption("repos"))) 
library(microViz)
library(phyloseq)

#order ps based on how we want facet layout
ps@sam_data$sample_type_depth <- factor(ps@sam_data$sample_type_depth, levels = c("sediment_deep","sediment_surface", "peat_deep", "peat_surface", "soil_surface"))
# order samples by latitude in each substrate
ps <- ps %>% ps_arrange(sample_type_depth, desc(latitude),sample_id)
ps_tr <- transform_sample_counts(ps, function(x) x / sum(x) *100) #transform ps OTU level

```

# Richness
```{r}
#richness <- estimate_richness(ps)
head(richness)
saveRDS(richness, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/richness.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/richness.rds")

cols<-c("orange","#fbd1a1", "dodgerblue","#A1C7E0", "#ef767a")

plot_richness(ps, x="sample_id", color="sample_type_depth", measures=c("Observed"))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(size=2))+
    scale_color_manual(values=cols) 

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Alpha_diversity",variation,".png"), height = 7, width=9, units="in")

rm(richness)
```

# NMDS
```{r}
set.seed(484)
nmds <- ps %>% ordinate(., method = "NMDS", distance="bray")
round(nmds$stress,digits=3)

saveRDS(nmds, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/nmds.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/nmds.rds")

```

#### Plot NMDS
```{r}
library(remotes)
nmds_points <- as.data.frame(nmds$points) %>% rownames_to_column(var="sample_id")
nmds_meta<-merge(nmds_points,metadata, by="sample_id")

environmental.envfit<- envfit(nmds ~latitude+ longitude+ elevation + measured_temperature+Tann+ Tmin+ Tmax+ MAP, data=metadata, perm=1000,na.rm=T)   

env.vectors <- as.data.frame(scores(environmental.envfit, display = "vectors"))
# * ordiArrowMul(environmental.envfit)
env.vectors.pval <- env.vectors %>% mutate(p=environmental.envfit$vectors$pvals,    Variable=row.names(env.vectors))  %>% 
        filter(p<0.05) 


cols<-c("orange","#fbd1a1", "dodgerblue","#A1C7E0", "#ef767a")

ggplot()+
  geom_point(data=nmds_meta, aes(x=MDS1, y=MDS2, color=sample_type_depth))+
  geom_segment(data = env.vectors.pval, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
        size =1, alpha = 0.5, colour = "black", arrow = arrow(length=unit(1, "mm")), lwd=0.5) +
  geom_text(data = env.vectors.pval, aes(x = NMDS1, y = NMDS2, label = Variable),
            size = 3)+
  custom_theme() +
  scale_color_manual(values=cols) + 
  labs(x = "NMDS1", y = "NMDS2", title="OTU level",
             color= "Substrate",
             size="Depth (cm)") +   
  annotate("text", x = min(nmds_meta$MDS1), y = max(nmds_meta$MDS2), 
           label = paste("Stress =", round(nmds$stress,digits=3)), 
           hjust = 0, vjust = 1, size = 4) 

ggsave(device = "png",filename=paste0("../5_Figures/FigX_NMDS_OTU_envfit",variation,".png"), height = 7, width=9, units="in")
```

# SKIP Merge samples by site: cant merge by site because now we have surface and deep samples 
```{r}
#create another column in the metadata with factored sample_type. This is because sample type and sample id disappears when summing samples within sites.
sample_type.ps<- data.frame(ps@sam_data$sample_type_depth)
colnames(sample_type.ps)
sample_type.ps<-  sample_type.ps %>% mutate(sample_type.factored = as.numeric(factor(ps.sam_data.sample_type_depth)))


#Add sample_type_factored column to sample data 
sample_data(ps)$sample_type_factored <- as.integer(sample_type.ps$sample_type.factored)
# Merge samples by site 
ps.merged.by.site<- merge_samples(ps, "site_name") # summed by site
#add back site name in column
sample_data(ps.merged.by.site)$site_name_relabeled <-rownames(ps.merged.by.site@sam_data)
sample_data(ps.merged.by.site)$sample_type_factored <-as.integer(ps.merged.by.site@sam_data$sample_type_factored)

#get factored "sample_type" and assign character then add it to sample_data of merged samples
to_change<- as.data.frame(ps.merged.by.site@sam_data$sample_type_factored)
colnames(to_change)<- "sample_type_factored"
to_change<- to_change %>%
  mutate(sample_type_merged = case_when(sample_type_factored==1 ~ "peat surface",
                                        sample_type_factored==2 ~ "peat deep",
                                       sample_type_factored==3 ~ "sediment surface",
                                       sample_type_factored==4 ~ "sediment deep",
                                       sample_type_factored==5 ~"soil"))

sample_data(ps.merged.by.site)$sample_type_merged <-to_change$sample_type_merged

ps.merged.by.site.samp.dat=data.frame(ps.merged.by.site@sam_data)

# transform counts
ps_tr.merged.by.site <- transform_sample_counts(ps.merged.by.site, function(x) x / sum(x))


```


# Plot kingdom and phyla
```{r}
ps.kingdom <- tax_glom(ps_tr, "Kingdom")

library(rcartocolor)
nColor <- 2
scales::show_col(carto_pal(nColor, "Safe"))

plot_bar(ps.kingdom, fill = "Kingdom", title = "") + 
  geom_bar(aes( fill=Kingdom, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=1))+ labs(x="", y="% taxa")
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Kingdom_merged_by_site_",variation,".png"), height = 7, width=15, units="in")


# Gloom by phylum
ps.tr_phylum= tax_glom(ps_tr, "ps.tr_phylum")
saveRDS(ps_tr.common.ps, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.tr_phylum.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.tr_phylum.rds")


cols <- c("plum", "lightskyblue", "orange", "darkolivegreen1", "chocolate4", "dodgerblue","firebrick1", "bisque3", "gold4", "darkseagreen1", "purple3", "mediumvioletred", "forestgreen", "khaki1", "hotpink",
"mediumblue", "#6DDE88", "aquamarine", "lightpink1", "cornsilk", "grey50", "darkorchid1", 
"lightblue", "darkorange" )

color.vector <- colorRampPalette(cols)(84) 
mycols<- sample(color.vector)
pie(rep(1, 84), col = mycols) #randomize the colors


plot_bar(ps.tr_phylum, fill = "Phylum", title = "") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="% taxa")+
  scale_x_discrete()+
  scale_fill_manual( values = mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Phylum",variation,".png"), height = 7, width=15, units="in")

```


# Get shared and not shared taxa 
```{r}
# get taxa  common across all 3 environments
ps_tr.common.taxa<- common_taxa(ps_tr, treatment="sample_type", n = "all")
ps_tr.common.subset <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% ps_tr.common.taxa)
ps_tr.common.ps <- merge_phyloseq(ps_tr.common.subset, tax_table(ps_tr), sample_data(ps_tr))
ps_tr.common.ps<-tax_glom(ps_tr.common.ps, "Phylum")
saveRDS(ps_tr.common.ps, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.rds")

# get taxa not common across all 3 environments
notcommon.taxa<-rownames(otu_table(ps_tr))[! rownames(otu_table(ps_tr)) %in% ps_tr.common.taxa]
ps_tr.notcommon.subset <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% notcommon.taxa)
ps_tr.notcommon.ps <- merge_phyloseq(ps_tr.notcommon.subset, tax_table(ps_tr), sample_data(ps_tr))
ps_tr.notcommon.ps<-tax_glom(ps_tr.notcommon.ps, "Phylum")
saveRDS(ps_tr.notcommon.ps, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.notcommon.ps.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.notcommon.ps.rds")

# get taxa  common across all 5 environments
ps_tr.common.taxa.5<- common_taxa(ps_tr, treatment="sample_type_depth", n = "all")
ps_tr.common.subset.5 <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% ps_tr.common.taxa.5)
ps_tr.common.ps.5 <- merge_phyloseq(ps_tr.common.subset.5, tax_table(ps_tr), sample_data(ps_tr))
ps_tr.common.ps.5<-tax_glom(ps_tr.common.ps.5, "Phylum")
saveRDS(ps_tr.common.ps.5, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.5.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.5.rds")

# get taxa not common across all 5 environments
notcommon.taxa.5<-rownames(otu_table(ps_tr))[! rownames(otu_table(ps_tr)) %in% ps_tr.common.taxa.5]
ps_tr.notcommon.subset.5 <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% notcommon.taxa.5)
ps_tr.notcommon.ps.5 <- merge_phyloseq(ps_tr.notcommon.subset.5, tax_table(ps_tr), sample_data(ps_tr))
ps_tr.notcommon.ps.5<-tax_glom(ps_tr.notcommon.ps.5, "Phylum")
saveRDS(ps_tr.common.ps.5, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.5.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps_tr.common.ps.5.rds")

```

###  Plot shared & not shared taxa 
```{r}
plot_bar(ps_tr.common.ps, fill = "Phylum", title = "Shared taxa across 3 substrates at OTU level (sediment, peat, soil) ") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="% taxa")+
  scale_x_discrete() + 
  scale_y_continuous(limits=c(0,100))+
  scale_fill_manual( values = mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Shared_phylum_3_substrates",variation,".png"), height = 7, width=15, units="in")


plot_bar(ps_tr.notcommon.ps, fill = "Phylum", title = "Taxa not shared across 3 substrates at OTU level (sediment, peat, soil)") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~ sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="% taxa")+
  scale_x_discrete()+
  scale_y_continuous(limits=c(0,100))+
  scale_fill_manual( values = mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_NotShared_phylum_3_substrates",variation,".png"), height = 7, width=15, units="in")

plot_bar(ps_tr.common.ps.5, fill = "Phylum", title = "Shared taxa across all 5 substrates at OTU level") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="% taxa")+
  scale_x_discrete() + 
  scale_y_continuous(limits=c(0,100))+
  scale_fill_manual( values = mycols)


ggsave(device = "png",filename=paste0("../5_Figures/FigX_Shared_phylum_5_substrates",variation,".png"), height = 7, width=15, units="in")


plot_bar(ps_tr.notcommon.ps.5, fill = "Phylum", title = "Taxa not shared across all 5 substrates at OTU level") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~ sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="% taxa")+
  scale_x_discrete()+
  scale_y_continuous(limits=c(0,100))+
  scale_fill_manual( values = mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_NotShared_phylum_5_substrates",variation,".png"), height = 7, width=15, units="in")

```

# Get unique taxa 
```{r}
unq_taxa.3<- unique_taxa(ps_tr,  treatment ="sample_type", subset = NULL)
unq_taxa_combined.3 = unlist(unq_taxa.3, use.names=FALSE)
ps_unq_subset.3 <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% unq_taxa_combined.3)
ps.3.unq <- merge_phyloseq(ps_unq_subset.3, tax_table(ps_tr), sample_data(ps_tr))
ps.3.unq<-tax_glom(ps.3.unq, "Phylum")
saveRDS(ps.3.unq, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.3.unq.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.3.unq.rds")

rm(unq_taxa.3)
rm(unq_taxa_combined.3)
rm(t_otu_ps.3)
rm(ps_unq_subset.3)


unq_taxa.5<- unique_taxa(ps_tr,  treatment ="sample_type_depth", subset = NULL)
unq_taxa_combined.5 = unlist(unq_taxa.5, use.names=FALSE)
ps_unq_subset.5 <- subset(otu_table(ps_tr), rownames(otu_table(ps_tr)) %in% unq_taxa_combined.5)
ps.5.unq <- merge_phyloseq(ps_unq_subset.5, tax_table(ps_tr), sample_data(ps_tr))
ps.5.unq<-tax_glom(ps.5.unq, "Phylum")
saveRDS(ps.3.unq, "/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.5.unq.rds")
load("/Users/ioana/Desktop/MountainMicrobes/4_DataAnalysis/ps.5.unq.rds")

rm(unq_taxa.5)
rm(unq_taxa_combined.5)
rm(t_otu_ps.5)
rm(ps_unq_subset.5)
```


### Plot unique taxa 3 substrates
```{r}
plot_bar(ps.3.unq, fill = "Phylum", title = "Unique taxa 3 substrates") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="Fraction taxa")+
  scale_x_discrete() + scale_y_continuous(limits=c(0,100))+
  scale_fill_manual(values=mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Unique_taxa_by_phylum_3_substrates_",variation,".png"), height = 7, width=15, units="in")
```

### Plot unique taxa 5 substrates
```{r}
cols = rainbow(82, s=.6, v=.9)[sample(1:82,82)]

plot_bar(ps.5.unq, fill = "Phylum", title = "Unique taxa 5 substrates") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_depth, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=2))+ labs(x="", y="Fraction taxa")+
  scale_x_discrete() + scale_y_continuous(limits=c(0,100))+
  scale_fill_manual(values=mycols)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Unique_taxa_by_phylum_5_substrates_",variation,".png"), height = 7, width=15, units="in")
```


#Get top taxa PHYLUM levle 
```{r}
ps_tibb <-ps_tr %>%
  tax_glom("Phylum") %>%
  psmelt() %>%
  as_tibble()

# highest abundance: all samples within each environment pooled together
ps_tibb_average<-ps_tibb %>%
  group_by(sample_type_depth, Phylum) %>%
  summarise(mean_percent = mean(Abundance)) %>%
  arrange(sample_type_depth,-mean_percent) %>%
  #filter( mean_percent > 1) %>%
  print(n=60)

#write.csv(ps_tibb_average, "Top_phylum_5_substrates.csv")

ps_tibb_average<- read.csv("Top_phylum_5_substrates.csv")

#get top taxa that are common across all 5 substrates
ps_tr.common.taxa<-ps_tibb_average %>%
  filter( mean_percent > 0.5) %>%
  group_by(Phylum) %>%
  filter(n_distinct(sample_type_depth) == n_distinct(ps_tibb_average$sample_type_depth))  %>%
    arrange(sample_type_depth, Phylum)
  

ps_tr.common.taxa$sample_type_depth <- factor(ps_tr.common.taxa$sample_type_depth, levels = c("sediment_deep","sediment_surface", "peat_deep", "peat_surface", "soil_surface"))

ggplot(ps_tr.common.taxa,aes(as.character(Phylum),mean_percent, fill=sample_type_depth))+
        geom_bar(stat = "identity",color="black",size=0.2) +
        facet_wrap(~sample_type_depth,ncol=1)+
        scale_fill_manual(values = c("dodgerblue","#A1C7E0","orange","#fbd1a1",  "#ef767a")) +
   #     scale_y_continuous(limits=c(0,20), expand=c(0,0))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              axis.title.x=element_blank(),
              strip.background = element_blank(),
              strip.text.x = element_blank())+
        labs(y="Average % across all sites", title = "Top common PHYLA across all 5 substrates", fill="Substrate")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Top_Common_Phylum_5_substrates",variation,".png"), height = 8, width=10, units="in")


#get top 10 taxa within each of the 5 substrates
ps_tr.top.taxa<-ps_tibb_average %>%
  filter( mean_percent > 0) %>%
  group_by((sample_type_depth)) %>%
        slice_max(order_by = mean_percent , n = 10)

ps_tr.top.taxa$sample_type_depth <- factor(ps_tr.top.taxa$sample_type_depth, levels = c("sediment_deep","sediment_surface", "peat_deep", "peat_surface", "soil_surface"))


ggplot(ps_tr.top.taxa,aes(as.character(Phylum),mean_percent, fill=sample_type_depth))+
        geom_bar(stat = "identity",color="black",size=0.2) +
        facet_wrap(~sample_type_depth,ncol=1)+
        scale_fill_manual(values = c("dodgerblue","#A1C7E0","orange","#fbd1a1",  "#ef767a")) +
   #     scale_y_continuous(limits=c(0,20), expand=c(0,0))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              axis.title.x=element_blank(),
              strip.background = element_blank(),
              strip.text.x = element_blank())+
        labs(y="Average % across all sites", title = "Top 10 PHYLA within each of the 5 substrates", fill="Substrate")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Top_10_Phyla_in_each_substrate",variation,".png"), height = 8, width=10, units="in")

```
  
#Get top taxa FAMILY level
```{r}
ps_tibb <-ps_tr %>%
  tax_glom("Family") %>%
  psmelt() %>%
  as_tibble()

# highest abundance: all samples within each environment pooled together
ps_tibb_average<-ps_tibb %>%
  group_by(sample_type_depth, Family) %>%
  summarise(mean_percent = mean(Abundance)) %>%
  arrange(sample_type_depth,-mean_percent) %>%
  #filter( mean_percent > 1) %>%
  print(n=60)

#write.csv(ps_tibb_average, "Top_Family_5_substrates.csv")

ps_tibb_average<- read.csv("Top_Family_5_substrates.csv")

#get top taxa that are common across all 5 substrates
ps_tr.common.taxa<-ps_tibb_average %>%
  filter( mean_percent>0.1) %>%
  group_by(Family) %>%
  filter(n_distinct(sample_type_depth) == n_distinct(ps_tibb_average$sample_type_depth))  %>%
    arrange(sample_type_depth, Family)
  

ps_tr.common.taxa$sample_type_depth <- factor(ps_tr.common.taxa$sample_type_depth, levels = c("sediment_deep","sediment_surface", "peat_deep", "peat_surface", "soil_surface"))

ggplot(ps_tr.common.taxa,aes(as.character(Family),mean_percent, fill=sample_type_depth))+
        geom_bar(stat = "identity",color="black",size=0.2) +
        facet_wrap(~sample_type_depth,ncol=1)+
        scale_fill_manual(values = c("dodgerblue","#A1C7E0","orange","#fbd1a1",  "#ef767a")) +
   #     scale_y_continuous(limits=c(0,20), expand=c(0,0))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              axis.title.x=element_blank(),
              strip.background = element_blank(),
              strip.text.x = element_blank())+
        labs(y="Average % across all sites", title = "Top common Family across all 5 substrates", fill="Substrate")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Top_Common_Family_5_substrates",variation,".png"), height = 8, width=10, units="in")


#get top 10 taxa within each of the 5 substrates
ps_tr.top.taxa<-ps_tibb_average %>%
  filter( mean_percent > 0.1) %>%
  group_by((sample_type_depth)) %>%
        slice_max(order_by = mean_percent , n = 10)

ps_tr.top.taxa$sample_type_depth <- factor(ps_tr.top.taxa$sample_type_depth, levels = c("sediment_deep","sediment_surface", "peat_deep", "peat_surface", "soil_surface"))


ggplot(ps_tr.top.taxa,aes(as.character(Family),mean_percent, fill=sample_type_depth))+
        geom_bar(stat = "identity",color="black",size=0.2) +
        facet_wrap(~sample_type_depth,ncol=1)+
        scale_fill_manual(values = c("dodgerblue","#A1C7E0","orange","#fbd1a1",  "#ef767a")) +
   #     scale_y_continuous(limits=c(0,20), expand=c(0,0))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
              axis.title.x=element_blank(),
              strip.background = element_blank(),
              strip.text.x = element_blank())+
        labs(y="Average % across all sites", title = "Top 10 Families within each of the 5 substrates", fill="Substrate")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Top_10_Families_in_each_substrate",variation,".png"), height = 8, width=10, units="in")
```



require("fantaxtic")
require("ggnested")
require("phyloseq")
top_asv <- top_taxa(ps, n_taxa = 20)

plot_nested_bar(ps_obj = top_asv$ps_obj,
                top_level = "Phylum",
                nested_level = "Species")+
        facet_grid(~sample_type_depth, scales="free") 


#Run DESeq2
```{r}
#http://joey711.github.io/phyloseq-extensions/DESeq2.html
library("DESeq2")
packageVersion("DESeq2")
ps_plus1<-ps.merged.by.site
otu_table(ps_plus1)<- otu_table(ps.merged.by.site) + 1 #because there are 0 counts and can't compute geometric mean below

diagdds1 = phyloseq_to_deseq2(ps_plus1, ~sample_type_merged)
diagdds2 = DESeq(diagdds1, test="Wald", fitType="parametric")

#investigate results
res = results(diagdds2, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps_plus1)[rownames(sigtab), ],"matrix"))
head(sigtab)
```
### Plot DESeq2 
```{r}
theme_set(theme_bw())
scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)}

# Kingdom order
x = tapply(sigtab$log2FoldChange, sigtab$Kingdom, function(x) max(x))
x = sort(x, TRUE)
sigtab$Kingdom = factor(as.character(sigtab$Kingdom), levels=names(x))
    
# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))

# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))

ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + 
  geom_point(size=2) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5),
  axis.text.x.bottom = element_text(size=5))+
   guides(color=guide_legend(ncol=2))+
        labs(title="p<0.01")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Statistically_sig_genus",variation,".png"), height = 8, width=15, units="in")

ggplot(sigtab, aes(x=Phylum, y=log2FoldChange, color=Phylum)) + 
  geom_point(size=2) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5),
  axis.text.x.bottom = element_text(size=5))+
   guides(color=guide_legend(ncol=2))+
                labs(title="p<0.01")


ggsave(device = "png",filename=paste0("../5_Figures/FigX_Statistically_sig_phylum",variation,".png"), height = 8, width=15, units="in")

```


## Calculate community similarity
```{r}
# used ps object because normalized and sums to one
OTU<-t(as.data.frame(ps@otu_table@.Data))
#calculate Bray-Curtis similarity between all samples
comm.dist <- 1 - vegdist(OTU)

#convert the dissimilarity matrix to long format
comm.dist.ls<-as.matrix(comm.dist)
comm.dist.ls[upper.tri(comm.dist.ls, diag = T)] <- NA
comm.dist.ls<-reshape2::melt(comm.dist.ls, na.rm=T)
names(comm.dist.ls)<-c("s1_sample_id","s2_sample_id","comm")

#merge in metadata to table
meta_sub <- metadata %>% select(sample_id, sample_type, latitude, longitude) %>% na.omit(.)
metadata_s1<-meta_sub
colnames(metadata_s1)<-paste("s1",colnames(metadata_s1), sep="_")
metadata_s2<-meta_sub
colnames(metadata_s2)<-paste("s2",colnames(metadata_s2), sep="_")

# convert to dataframe, change the sample IDs from factors to characters and merge in the metadata tables
comps<-data.frame(comm.dist.ls)
names(comps)<-c("s1_sample_id","s2_sample_id","comm")

comps$s1_sample_id<-as.character(comps$s1_sample_id)
comps$s2_sample_id<-as.character(comps$s2_sample_id)
comps<-merge(comps, metadata_s1, by="s1_sample_id")
comps<-merge(comps, metadata_s2, by="s2_sample_id")


#remove metadata notes
comps$s1_notes<-NULL
comps$s2_notes<-NULL

comps <- comps %>% mutate(sample_comparison = 
                  case_when(s1_sample_type==s2_sample_type ~ s1_sample_type,
                            s1_sample_type=="peat" & s2_sample_type=="soil" ~ "soil-peat",
                            s2_sample_type=="peat" & s1_sample_type=="soil" ~ "soil-peat",
                            s1_sample_type=="peat" & s2_sample_type=="sediment" ~ "peat-sediment",
                            s2_sample_type=="peat" & s1_sample_type=="sediment" ~ "peat-sediment",
                            s1_sample_type=="soil" & s2_sample_type=="sediment" ~ "soil-sediment",
                            s2_sample_type=="soil" & s1_sample_type=="sediment" ~ "soil-sediment")) %>% select(sample_comparison,comm)

comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))


rm(metadata_s1);rm(metadata_s2);rm(comm.dist.ls);rm(OTU); rm(ps_tr)
#write.csv(comps,paste0(Sys.Date(),"_community_similarity_sample_type.csv"))
```

### Plot community similarity
```{r}
comps<-read.csv("2024-08-06_community_similarity_sample_type.csv")
comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))


colors <- c("soil"="#A6761D", 
                 "peat"="#D95F02", 
                 "sediment"="#7570B3", 
                 "soil-peat"="#E7298A", 
                 "peat-sediment"="#66A61E", 
                 "soil-sediment"="#E6AB02")

ggplot(data=comps, aes(y=comm, x=sample_comparison, fill=sample_comparison)) + 
        geom_boxplot() + custom_theme() + 
        labs(x="",
             y="Community similiary (1 - Bray)",
             fill="") +
        scale_fill_manual(values=colors)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_sample_type_",variation,".png"), height = 7, width=10, units="in")

```


## Geographic distance
some of Dulcinea's peat samples don't have lat/lon, so this code has been made to remove those samples. need to update this after values have been updated 
```{r}
#subset data by those with lat/longs
meta_sub <- metadata %>% select(sample_id, sample_type, latitude, longitude) %>% na.omit(.)

# used ps object because normalized and sums to one
OTU<-t(as.data.frame(ps@otu_table@.Data))

#subset samples that have lat/long
OTU<-OTU[rownames(meta_sub),]
table(rownames(OTU)==rownames(meta_sub)) #all true
table(rownames(meta_sub)==rownames(OTU)) #all true

#calculate Bray-Curtis similarity between all samples
comm.dist <- 1 - vegdist(OTU)


# Lat&lon to distance in meters
xy <- data.frame(X = meta_sub$longitude, Y = meta_sub$latitude)
rownames(xy)<-meta_sub$sample_id
dist_m_output<-distm(xy)
rownames(dist_m_output)<-rownames(xy)
names(dist_m_output)<-rownames(xy)
dist_m_output<-as.dist(dist_m_output)


#transform all distance matrices into dataframes with pairwise comparisons
coord.dist.ls<-as.matrix(dist_m_output)
coord.dist.ls[upper.tri(coord.dist.ls, diag = T)] <- NA
coord.dist.ls<-reshape2::melt(coord.dist.ls, na.rm=T)
names(coord.dist.ls)<-c("s1_sample_id","s2_sample_id","geo_dist")
coord.dist.ls$s1_sample_id<-as.character(coord.dist.ls$s1_sample_id)
coord.dist.ls$s2_sample_id<-as.character(coord.dist.ls$s2_sample_id)

# convert community distances to long format
comm.dist.ls<-as.matrix(comm.dist)
comm.dist.ls[upper.tri(comm.dist.ls, diag = T)] <- NA
comm.dist.ls<-reshape2::melt(comm.dist.ls, na.rm=T)
names(comm.dist.ls)<-c("s1_sample_id","s2_sample_id","comm")
comm.dist.ls$s1_sample_id<-as.character(comm.dist.ls$s1_sample_id)
comm.dist.ls$s2_sample_id<-as.character(comm.dist.ls$s2_sample_id)

#check the names of these match
table(coord.dist.ls[,1]==comm.dist.ls[,1]) #all true
table(coord.dist.ls[,2]==comm.dist.ls[,2]) #all true

#create df with similarity of community and distance
comps<-data.frame(comm.dist.ls)
comps$geo_dist<-coord.dist.ls$geo_dist

#merge in metadata to table
metadata_s1<-meta_sub
colnames(metadata_s1)<-paste("s1",colnames(metadata_s1), sep="_")

metadata_s2<-meta_sub
colnames(metadata_s2)<-paste("s2",colnames(metadata_s2), sep="_")

comps<-merge(comps, metadata_s1, by="s1_sample_id")
comps<-merge(comps, metadata_s2, by="s2_sample_id")

#remove metadata notes
comps$s1_notes<-NULL
comps$s2_notes<-NULL

# convert to km and add in sample comparison categories
comps <- comps %>% mutate(geo_dist_km=geo_dist/1000) %>%
        mutate(sample_comparison = 
                  case_when(s1_sample_type==s2_sample_type ~ s1_sample_type,
                            s1_sample_type=="peat" & s2_sample_type=="soil" ~ "soil-peat",
                            s2_sample_type=="peat" & s1_sample_type=="soil" ~ "soil-peat",
                            s1_sample_type=="peat" & s2_sample_type=="sediment" ~ "peat-sediment",
                            s2_sample_type=="peat" & s1_sample_type=="sediment" ~ "peat-sediment",
                            s1_sample_type=="soil" & s2_sample_type=="sediment" ~ "soil-sediment",
                            s2_sample_type=="soil" & s1_sample_type=="sediment" ~ "soil-sediment")) %>% select(sample_comparison,comm, geo_dist_km)

comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))

 write.csv(comps,paste0(Sys.Date(),"_community_similarity_sample_type_geodist.csv"))

```


### Plot geographic distance with community similarity
```{r}
comps<-read.csv("2024-07-13_community_similarity_sample_type_geodist.csv")
comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))
colors <- c("soil"="#A6761D", 
                 "peat"="#D95F02", 
                 "sediment"="#7570B3", 
                 "soil-peat"="#E7298A", 
                 "peat-sediment"="#66A61E", 
                 "soil-sediment"="#E6AB02")

ggplot(data=comps, aes(y=comm, x=geo_dist_km, color=sample_comparison)) + 
        geom_point() + custom_theme() + 
        labs(x="Geographic distance (km)",
             y="Community similiary (1 - Bray)",
             color="")+
        scale_color_manual(values=adjustcolor(colors, alpha.f = 0.5))

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_sample_type_geo_distance",variation,".png"), height = 7, width=9, units="in")

ggplot(data=comps, aes(y=comm, x=geo_dist_km)) + 
        geom_point() + custom_theme() + 
        labs(x="Geographic distance (km)",
             y="Community similiary (1 - Bray)")+ 
        geom_smooth(method='lm', formula= y~x, col="palegreen3")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_geo_distance_lm_",variation,".png"), height = 7, width=9, units="in")

```

# check this analysis
```{r}
Check this one for mean abundace to rank it
https://github.com/joey711/phyloseq/issues/631
clusterData = psmelt(ps_tr)
clusterData = filter(clusterData,Abundance > 0)

# this is where the mean is calculated and the taxa to display is chosen
clusterAgg = aggregate(Abundance ~ OTU + phylum,data=clusterData,mean)

# filtering and picking the number to display
clusterAgg = clusterAgg[order(-clusterAgg$Abundance),][1:100,]

ggplot(clusterAgg,aes(x=reorder(OTU,-Abundance),y=Abundance)) +
    geom_point(aes(color=phylum),size=3) + 
    theme(axis.ticks = element_blank(), axis.text.x = element_blank()) +
    scale_y_log10()
rm(ps_tr.merged.by.site)
```


# Switch to GENUS and rerun all
```{r}
source("../3_CreatePhyloseq/3_CreatePhyloseq_ZOTUexact.R")

ps_genus<-tax_glom(ps, "Genus")
saveRDS(ps_genus, "ps_genus.rds")
#load("ps_genus.rds")

#ps<-ps_genus
#ps_tr<- transform_sample_counts(ps, function(x) x / sum(x) *100) #transform ps Genus level
#variation="ZOTUexact_GENUSlevel"

# rerun desired analyses at genus level
```


# Shared ESVs - not started
```{r}
#calculate
shared_esvs <- phyloseq_num_shared_otus(ps)

#shared ESVs
shared<-as.matrix(shared_esvs[["shared"]])
shared[upper.tri(shared, diag = T)] <- NA
shared<-melt(shared,na.rm=TRUE)

colnames(shared)[3]<-"shared"

#nonshared ESVs
nonshared<-as.matrix(shared_esvs[["nonshared_total"]])
nonshared[upper.tri(nonshared, diag = T)] <- NA
nonshared<-melt(nonshared,na.rm=TRUE)
colnames(nonshared)[3]<-"nonshared"

table(nonshared$Var1==shared$Var1) #all T
table(nonshared$Var2==shared$Var2) #all T

shared$nonshared<-nonshared$nonshared
rm(nonshared)
shared$percent<-shared$shared/(shared$shared+shared$nonshared)
```

