---
title: "16S Microbial Analysis from lake sediments, peat, and soils in Wyoming mountains"
author: "Created by Jordan Von Eggers, modified by Ioana Stefanscu"
date: "`r Sys.Date()`"
output: pdf_document
---

# Load packages and report session info
```{r}
require(phyloseq)
require(tidyverse)
require(vegan)

# newly added packages that may need installing 
require(reshape2)
require(metagMisc)
require(RColorBrewer)
require(geosphere)

# #installation for metagMisc
# install.packages("remotes")
# remotes::install_github("vmikk/metagMisc")

sessionInfo()
```


R version 4.4.0 (2024-04-24)

other attached packages:
 [1] vegan_2.6-6.1   lattice_0.22-6  permute_0.9-7   lubridate_1.9.3 forcats_1.0.0   stringr_1.5.1   dplyr_1.1.4     purrr_1.0.2     readr_2.1.5    
[10] tidyr_1.3.1     tibble_3.2.1    ggplot2_3.5.1   tidyverse_2.0.0 phyloseq_1.48.0


# Load data
(this takes a few minutes)
```{r}
source("../3_CreatePhyloseq/3_CreatePhyloseq_ZOTUexact.R")
```

# Custom theme
```{r}
  custom_theme <- function() {
  theme_bw() +
    theme(
      text = element_text(color = "black", size = 13),
      axis.text = element_text(color = "black",size = 13),
      axis.title = element_text(color = "black",size = 13),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      legend.text = element_text(size = 13),
      legend.title = element_text(size = 13),
      strip.text = element_text(size = 13, color = "black"))}
```

#Remove samples ???
```{r}
# to_remove <- c("Sample1", "Sample5", "Sample10", "Sample30")
# ps<- prune_samples(!(sample_names(ps) %in% to_remove), ps)
```



#Richness
```{r}
richness <- estimate_richness(ps)
head(richness)

plot_richness(ps, x="sample_id", color="sample_type", measures=c("Observed"))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(size=2))


ggsave(device = "png",filename=paste0("../5_Figures/FigX_Alpha_diversity",variation,".png"), height = 7, width=9, units="in")
        
```

# NMDS
```{r}
set.seed(484)
nmds <- ps %>% ordinate(., method = "NMDS", distance="bray")
round(nmds$stress,digits=3)
```
Great stress score: 0.1388955!
Updated score:  0.1394602  Ioana Stefanescu, 07.09.2024

```{r}
nmds_points <- as.data.frame(nmds$points) %>% rownames_to_column(var="sample_id")
nmds_meta<-merge(nmds_points,metadata,by="sample_id")

rm(nmds_points)


environmental.envfit= envfit(nmds ~latitude+ longitude+ elevation + depth_start+
depth_end+ Tann+ Tmin+ Tmax+ MAP, data=metadata, perm=1000,na.rm=T)   

env.vectors <- as.data.frame(scores(environmental.envfit, display = "vectors"))* ordiArrowMul(environmental.envfit)
env.vectors.pval <- env.vectors %>% mutate(p=environmental.envfit$vectors$pvals,    Variable=row.names(env.vectors))  %>% 
        filter(p<0.05) 



cols<-c("#fbd1a2", "#A1C7E0", "#ef767a")

ggplot()+
        geom_point(data=nmds_meta, aes(x=MDS1, y=MDS2, color=sample_type, size=depth_start))+
        scale_color_manual(values=cols) + 
        custom_theme() +
        labs(x = "NMDS1",
             y = "NMDS2",
             color= "Environment",
             size="Depth (cm)") +   
        annotate("text", x = min(nmds_meta$MDS1), y = max(nmds_meta$MDS2), 
           label = paste("Stress =", round(nmds$stress,digits=3)), 
           hjust = 0, vjust = 1, size = 5) +
 geom_segment(data = env.vectors.pval, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
        size =1, alpha = 0.5, colour = "black", arrow = arrow(length=unit(2, "mm")), lwd=0.5) +
 geom_text(data = env.vectors.pval, aes(x = NMDS1, y = NMDS2, label = Variable),
            size = 5)


ggsave(device = "png",filename=paste0("../5_Figures/FigX_NMDS_sample_type_envfit",variation,".png"), height = 7, width=9, units="in")
```


# Merge samples by site
```{r}
#create another column in the metadata with factored sample_type. This is because sample type and sample id disappears when summing samples within sites.
sample_type.ps<- data.frame(ps@sam_data$sample_type)
colnames(sample_type.ps)
sample_type.ps<-  sample_type.ps %>% mutate(sample_type.factored = as.numeric(factor(ps.sam_data.sample_type)))


#Add sample_type_factored column to sample data 
sample_data(ps)$sample_type_factored <- as.integer(sample_type.ps$sample_type.factored)
# Merge samples by site 
ps.merged.by.site<- merge_samples(ps, "site_name") # summed by site
#add back site name in column
sample_data(ps.merged.by.site)$site_name_relabeled <-rownames(ps.merged.by.site@sam_data)

#get factored "sample_type" and assign character then add it to sample_data of merged samples
to_change<- as.data.frame(ps.merged.by.site@sam_data$sample_type_factored)
colnames(to_change)<- "sample_type_factored"
to_change<- to_change %>%
  mutate(sample_type_merged = case_when(sample_type_factored==1 ~ "peat",
                                       sample_type_factored==2 ~ "sediment",
                                       sample_type_factored==3~"soil"))

sample_data(ps.merged.by.site)$sample_type_merged <-to_change$sample_type_merged

ps.merged.by.site.samp.dat=data.frame(ps.merged.by.site@sam_data)

# transform counts
ps_tr.merged.by.site <- transform_sample_counts(ps.merged.by.site, function(x) x / sum(x))
# Gloom by kingdom
ps.merged.by.site.kingdom= tax_glom(ps_tr.merged.by.site, "Kingdom")
# Gloom by phylum
ps.merged.by.site.phylum= tax_glom(ps_tr.merged.by.site, "Phylum")



plot_bar(ps.merged.by.site.phylum, fill = "Kingdom", title = "Merged samples by site") + 
  geom_bar(aes( fill=Kingdom, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=1))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Kingdom_merged_by_site_",variation,".png"), height = 7, width=9, units="in")


plot_bar(ps.merged.by.site.phylum, fill = "Phylum", title = "Merged samples by site") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=4))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Phylum_merged_by_site_",variation,".png"), height = 7, width=15, units="in")

```


# Get common taxa across environments
```{r}
#remotes::install_github('schuyler-smith/phylosmith')
library(phylosmith)
merged.common.taxa<- common_taxa(ps_tr.merged.by.site, treatment="sample_type_merged", n = "all")

ps.merged.common.subset <- subset(otu_table(ps_tr.merged.by.site), select = colnames(otu_table(ps_tr.merged.by.site)) %in% merged.common.taxa)
ps.merged.common <- merge_phyloseq(ps.merged.common.subset, tax_table(ps_tr.merged.by.site), sample_data(ps_tr.merged.by.site))


plot_bar(ps.merged.common, fill = "Phylum", title = "Shared taxa by site") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=1))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Shared_phylum_by_site_",variation,".png"), height = 7, width=15, units="in")

```


#Get top taxa by environment  #### needs work, not finished #####
```{r}
data.ps <-ps_tr %>%
  tax_glom("Phylum") %>%
  psmelt() %>%
  as_tibble()

# highest abundance: all samples within each environment pooled together
data.ps %>%
  group_by(Phylum) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(sample_type) %>%
  arrange(-Abundance)%>%
  print(n=15)

```




# Notes for meeting June 27, 2024:
- May be worth removing samples that we know won't be in the analysis and repeat 1_RawSequenceProcessing step:
        - MAWC possblank samples, Calder not top and bottom sediments, anything else? (DONE)
- Need to select the correct samples and rerun the ZOTUexact 2_DataCleaning step (not done)
- May be worth looking at the ZOTU99, OTU99, OTU97 for 2_DataCleaning (not done)





# 27 June 2024 - Jordan Updating NMDS
# July 9&10, 2024 - Ioana Updating NMDS
# July 12, 2024 - Ioana Updating NMDS

# Community similarity - July 13, 2024 - Jordan
July 13, 2024 Jordan looking at community similarity within each sample type and across the other sample types. 

## Calculate community similarity
```{r}
# used ps object because normalized and sums to one
OTU<-t(as.data.frame(ps@otu_table@.Data))
#calculate Bray-Curtis similarity between all samples
comm.dist <- 1 - vegdist(OTU)

#convert the dissimilarity matrix to long format
comm.dist.ls<-as.matrix(comm.dist)
comm.dist.ls[upper.tri(comm.dist.ls, diag = T)] <- NA
comm.dist.ls<-reshape2::melt(comm.dist.ls, na.rm=T)
names(comm.dist.ls)<-c("s1_sample_id","s2_sample_id","comm")

#merge in metadata to table
metadata_s1<-metadata
colnames(metadata_s1)<-paste("s1",colnames(metadata_s1), sep="_")
metadata_s2<-metadata
colnames(metadata_s2)<-paste("s2",colnames(metadata_s2), sep="_")

# convert to dataframe, change the sample IDs from factors to characters and merge in the metadata tables
comps<-data.frame(comm.dist.ls)
comps$s1_sample_id<-as.character(comps$s1_sample_id)
comps$s2_sample_id<-as.character(comps$s2_sample_id)
comps<-merge(comps, metadata_s1, by="s1_sample_id")
comps<-merge(comps, metadata_s2, by="s2_sample_id")


#remove metadata notes
comps$s1_notes<-NULL
comps$s2_notes<-NULL

comps <- comps %>% mutate(sample_comparison = 
                  case_when(s1_sample_type==s2_sample_type ~ s1_sample_type,
                            s1_sample_type=="peat" & s2_sample_type=="soil" ~ "soil-peat",
                            s2_sample_type=="peat" & s1_sample_type=="soil" ~ "soil-peat",
                            s1_sample_type=="peat" & s2_sample_type=="sediment" ~ "peat-sediment",
                            s2_sample_type=="peat" & s1_sample_type=="sediment" ~ "peat-sediment",
                            s1_sample_type=="soil" & s2_sample_type=="sediment" ~ "soil-sediment",
                            s2_sample_type=="soil" & s1_sample_type=="sediment" ~ "soil-sediment")) %>% select(sample_comparison,comm)

comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))


rm(metadata_s1);rm(metadata_s2);rm(comm.dist.ls);rm(OTU); rm(ps_tr)
# write.csv(comps,paste0(Sys.Date(),"_community_similarity_sample_type.csv"))
```

### Plot community similarity
```{r}
comps<-read.csv("2024-07-13_community_similarity_sample_type.csv")
comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))


colors <- c("soil"="#A6761D", 
                 "peat"="#D95F02", 
                 "sediment"="#7570B3", 
                 "soil-peat"="#E7298A", 
                 "peat-sediment"="#66A61E", 
                 "soil-sediment"="#E6AB02")

ggplot(data=comps, aes(y=comm, x=sample_comparison, fill=sample_comparison)) + 
        geom_boxplot() + custom_theme() + 
        labs(x="",
             y="Community similiary (1 - Bray)",
             fill="") +
        scale_fill_manual(values=colors)

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_sample_type_",variation,".png"), height = 7, width=10, units="in")

```


## Geographic distance
some of Dulcinea's peat samples don't have lat/lon, so this code has been made to remove those samples. need to update this after values have been updated 
```{r}
#subset data by those with lat/longs
meta_sub <- metadata %>% select(sample_id, sample_type, latitude, longitude) %>% na.omit(.)

# used ps object because normalized and sums to one
OTU<-t(as.data.frame(ps@otu_table@.Data))

#subset samples that have lat/long
OTU<-OTU[rownames(meta_sub),]
table(rownames(OTU)==rownames(meta_sub)) #all true
table(rownames(meta_sub)==rownames(OTU)) #all true

#calculate Bray-Curtis similarity between all samples
comm.dist <- 1 - vegdist(OTU)


# Lat&lon to distance in meters
xy <- data.frame(X = meta_sub$longitude, Y = meta_sub$latitude)
rownames(xy)<-meta_sub$sample_id
dist_m_output<-distm(xy)
rownames(dist_m_output)<-rownames(xy)
names(dist_m_output)<-rownames(xy)
dist_m_output<-as.dist(dist_m_output)


#transform all distance matrices into dataframes with pairwise comparisons
coord.dist.ls<-as.matrix(dist_m_output)
coord.dist.ls[upper.tri(coord.dist.ls, diag = T)] <- NA
coord.dist.ls<-reshape2::melt(coord.dist.ls, na.rm=T)
names(coord.dist.ls)<-c("s1_sample_id","s2_sample_id","geo_dist")
coord.dist.ls$s1_sample_id<-as.character(coord.dist.ls$s1_sample_id)
coord.dist.ls$s2_sample_id<-as.character(coord.dist.ls$s2_sample_id)

# convert community distances to long format
comm.dist.ls<-as.matrix(comm.dist)
comm.dist.ls[upper.tri(comm.dist.ls, diag = T)] <- NA
comm.dist.ls<-reshape2::melt(comm.dist.ls, na.rm=T)
names(comm.dist.ls)<-c("s1_sample_id","s2_sample_id","comm")
comm.dist.ls$s1_sample_id<-as.character(comm.dist.ls$s1_sample_id)
comm.dist.ls$s2_sample_id<-as.character(comm.dist.ls$s2_sample_id)

#check the names of these match
table(coord.dist.ls[,1]==comm.dist.ls[,1]) #all true
table(coord.dist.ls[,2]==comm.dist.ls[,2]) #all true

#create df with similarity of community and distance
comps<-data.frame(comm.dist.ls)
comps$geo_dist<-coord.dist.ls$geo_dist

#merge in metadata to table
metadata_s1<-meta_sub
colnames(metadata_s1)<-paste("s1",colnames(metadata_s1), sep="_")

metadata_s2<-meta_sub
colnames(metadata_s2)<-paste("s2",colnames(metadata_s2), sep="_")

comps<-merge(comps, metadata_s1, by="s1_sample_id")
comps<-merge(comps, metadata_s2, by="s2_sample_id")

#remove metadata notes
comps$s1_notes<-NULL
comps$s2_notes<-NULL

# convert to km and add in sample comparison categories
comps <- comps %>% mutate(geo_dist_km=geo_dist/1000) %>%
        mutate(sample_comparison = 
                  case_when(s1_sample_type==s2_sample_type ~ s1_sample_type,
                            s1_sample_type=="peat" & s2_sample_type=="soil" ~ "soil-peat",
                            s2_sample_type=="peat" & s1_sample_type=="soil" ~ "soil-peat",
                            s1_sample_type=="peat" & s2_sample_type=="sediment" ~ "peat-sediment",
                            s2_sample_type=="peat" & s1_sample_type=="sediment" ~ "peat-sediment",
                            s1_sample_type=="soil" & s2_sample_type=="sediment" ~ "soil-sediment",
                            s2_sample_type=="soil" & s1_sample_type=="sediment" ~ "soil-sediment")) %>% select(sample_comparison,comm, geo_dist_km)

comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))

 write.csv(comps,paste0(Sys.Date(),"_community_similarity_sample_type_geodist.csv"))

```


### Plot geographic distance with community similarity
```{r}
comps<-read.csv("2024-07-13_community_similarity_sample_type_geodist.csv")
comps$sample_comparison<-factor(comps$sample_comparison, levels=c("soil","peat","sediment","soil-peat","peat-sediment","soil-sediment"))
colors <- c("soil"="#A6761D", 
                 "peat"="#D95F02", 
                 "sediment"="#7570B3", 
                 "soil-peat"="#E7298A", 
                 "peat-sediment"="#66A61E", 
                 "soil-sediment"="#E6AB02")

ggplot(data=comps, aes(y=comm, x=geo_dist_km, color=sample_comparison)) + 
        geom_point() + custom_theme() + 
        labs(x="Geographic distance (km)",
             y="Community similiary (1 - Bray)",
             color="")+
        scale_color_manual(values=adjustcolor(colors, alpha.f = 0.5))

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_sample_type_geo_distance",variation,".png"), height = 7, width=9, units="in")

ggplot(data=comps, aes(y=comm, x=geo_dist_km)) + 
        geom_point() + custom_theme() + 
        labs(x="Geographic distance (km)",
             y="Community similiary (1 - Bray)")+ 
        geom_smooth(method='lm', formula= y~x, col="palegreen3")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Community_similarity_geo_distance_lm_",variation,".png"), height = 7, width=9, units="in")

```



# Shared ESVs
```{r}
#calculate
shared_esvs <- phyloseq_num_shared_otus(ps)

#shared ESVs
shared<-as.matrix(shared_esvs[["shared"]])
shared[upper.tri(shared, diag = T)] <- NA
shared<-melt(shared,na.rm=TRUE)

colnames(shared)[3]<-"shared"

#nonshared ESVs
nonshared<-as.matrix(shared_esvs[["nonshared_total"]])
nonshared[upper.tri(nonshared, diag = T)] <- NA
nonshared<-melt(nonshared,na.rm=TRUE)
colnames(nonshared)[3]<-"nonshared"

table(nonshared$Var1==shared$Var1) #all T
table(nonshared$Var2==shared$Var2) #all T

shared$nonshared<-nonshared$nonshared
rm(nonshared)
shared$percent<-shared$shared/(shared$shared+shared$nonshared)
```


# Notes for meeting next, July 2024????
Do we want to merge samples by depth?
What other posibilities are there for merging samples?
 - JV comment: I think we should avoid merging samples because for peat and sediments, samples from the same core could be really different based off depth. If we want to show multiple abundances from one group, I think we should convert them to datapoints and plot them all to show the distribution of the data. Merging ends up losing some information in my opinion! the function psmelt() works great for this!
 - JV - "depth_start" for John's samples is the bottom centimeter interval - need to update this in the metadata. 

