---
title: "16S Microbial Analysis from lake sediments, peat, and soils in Wyoming mountains"
author: "Created by Jordan Von Eggers, modified by Ioana Stefanscu"
date: "`r Sys.Date()`"
output: pdf_document
---

# Load packages and report session info
```{r}
require(phyloseq)
require(tidyverse)
require(vegan)

sessionInfo()
```


R version 4.4.0 (2024-04-24)

other attached packages:
 [1] vegan_2.6-6.1   lattice_0.22-6  permute_0.9-7   lubridate_1.9.3 forcats_1.0.0   stringr_1.5.1   dplyr_1.1.4     purrr_1.0.2     readr_2.1.5    
[10] tidyr_1.3.1     tibble_3.2.1    ggplot2_3.5.1   tidyverse_2.0.0 phyloseq_1.48.0


# Load data
(this takes a few minutes)
```{r}
source("../3_CreatePhyloseq/3_CreatePhyloseq_ZOTUexact.R")
variation<-"ZOTUexact"
```

# Custom theme
```{r}
  custom_theme <- function() {
  theme_bw() +
    theme(
      text = element_text(color = "black", size = 13),
      axis.text = element_text(color = "black",size = 13),
      axis.title = element_text(color = "black",size = 13),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      legend.text = element_text(size = 13),
      legend.title = element_text(size = 13),
      strip.text = element_text(size = 13, color = "black"))}
```

#Remove samples ???
```{r}
# to_remove <- c("Sample1", "Sample5", "Sample10", "Sample30")
# ps<- prune_samples(!(sample_names(ps) %in% to_remove), ps)
```



#Richness
```{r}
richness <- estimate_richness(ps)
head(richness)

plot_richness(ps, x="sample_id", color="sample_type", measures=c("Observed"))+
        theme(panel.background = element_blank(),
              panel.border = element_rect(fill="transparent"),
              axis.text.x = element_text(size=2))


ggsave(device = "png",filename=paste0("../5_Figures/FigX_Alpha_diversity",variation,".png"), height = 7, width=9, units="in")
        
```

# NMDS
```{r}
set.seed(484)
nmds <- ps %>% ordinate(., method = "NMDS", distance="bray")
round(nmds$stress,digits=3)
```
Great stress score: 0.1388955!
Updated score:  0.1394602  Ioana Stefanescu, 07.09.2024

```{r}
nmds_points <- as.data.frame(nmds$points) %>% rownames_to_column(var="sample_id")
nmds_meta<-merge(nmds_points,metadata,by="sample_id")

rm(nmds_points)


environmental.envfit= envfit(nmds ~latitude+ longitude+ elevation + depth_start+
depth_end+ Tann+ Tmin+ Tmax+ MAP, data=metadata, perm=1000,na.rm=T)   

env.vectors <- as.data.frame(scores(environmental.envfit, display = "vectors"))* ordiArrowMul(environmental.envfit)
env.vectors.pval <- env.vectors %>% mutate(p=environmental.envfit$vectors$pvals,    Variable=row.names(env.vectors))  %>% 
        filter(p<0.05) 




cols<-c("#fbd1a2", "#A1C7E0", "#ef767a")

ggplot()+
        geom_point(data=nmds_meta, aes(x=MDS1, y=MDS2, color=sample_type, size=depth_start))+
        scale_color_manual(values=cols) + 
        custom_theme() +
        labs(x = "NMDS1",
             y = "NMDS2",
             color= "Environment",
             size="Depth (cm)") +   
        annotate("text", x = min(nmds_meta$MDS1), y = max(nmds_meta$MDS2), 
           label = paste("Stress =", round(nmds$stress,digits=3)), 
           hjust = 0, vjust = 1, size = 5) +
 geom_segment(data = env.vectors.pval, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
        size =1, alpha = 0.5, colour = "black", arrow = arrow(length=unit(2, "mm")), lwd=0.5) +
 geom_text(data = env.vectors.pval, aes(x = NMDS1, y = NMDS2, label = Variable),
            size = 5)


ggsave(device = "png",filename=paste0("../5_Figures/FigX_NMDS_sample_type_envfit",variation,".png"), height = 7, width=9, units="in")
```


# Merge samples by site
```{r}
#create another column in the metadata with factored sample_type. This is because sample type and sample id disappears when summing samples within sites.
sample_type.ps<- data.frame(ps@sam_data$sample_type)
colnames(sample_type.ps)
sample_type.ps<-  sample_type.ps %>% mutate(sample_type.factored = as.numeric(factor(ps.sam_data.sample_type)))


#Add sample_type_factored column to sample data 
sample_data(ps)$sample_type_factored <- as.integer(sample_type.ps$sample_type.factored)
# Merge samples by site 
ps.merged.by.site<- merge_samples(ps, "site_name") # summed by site
#add back site name in column
sample_data(ps.merged.by.site)$site_name_relabeled <-rownames(ps.merged.by.site@sam_data)

#get factored "sample_type" and assign character then add it to sample_data of merged samples
to_change<- as.data.frame(ps.merged.by.site@sam_data$sample_type_factored)
colnames(to_change)<- "sample_type_factored"
to_change<- to_change %>%
  mutate(sample_type_merged = case_when(sample_type_factored==1 ~ "peat",
                                       sample_type_factored==2 ~ "sediment",
                                       sample_type_factored==3~"soil"))

sample_data(ps.merged.by.site)$sample_type_merged <-to_change$sample_type_merged

ps.merged.by.site.samp.dat=data.frame(ps.merged.by.site@sam_data)

# transform counts
ps_tr.merged.by.site <- transform_sample_counts(ps.merged.by.site, function(x) x / sum(x))
# Gloom by kingdom
ps.merged.by.site.kingdom= tax_glom(ps_tr.merged.by.site, "Kingdom")
# Gloom by phylum
ps.merged.by.site.phylum= tax_glom(ps_tr.merged.by.site, "Phylum")



plot_bar(ps.merged.by.site.phylum, fill = "Kingdom", title = "Merged samples by site") + 
  geom_bar(aes( fill=Kingdom, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=1))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Kingdom_merged_by_site_",variation,".png"), height = 7, width=9, units="in")


plot_bar(ps.merged.by.site.phylum, fill = "Phylum", title = "Merged samples by site") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free")+
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=4))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Phylum_merged_by_site_",variation,".png"), height = 7, width=15, units="in")

```


# Get common taxa across environments
```{r}
#remotes::install_github('schuyler-smith/phylosmith')
library(phylosmith)
merged.common.taxa<- common_taxa(ps_tr.merged.by.site, treatment="sample_type_merged", n = "all")

ps.merged.common.subset <- subset(otu_table(ps_tr.merged.by.site), select = colnames(otu_table(ps_tr.merged.by.site)) %in% merged.common.taxa)
ps.merged.common <- merge_phyloseq(ps.merged.common.subset, tax_table(ps_tr.merged.by.site), sample_data(ps_tr.merged.by.site))


plot_bar(ps.merged.common, fill = "Phylum", title = "Shared taxa by site") + 
  geom_bar(aes( fill=Phylum, color=NULL), stat="identity", position="stack") + 
  facet_grid(~sample_type_merged, scales="free") +
  theme(plot.title = element_text(hjust=0.5, size=20), 
        panel.background = element_rect(color="black",fill="transparent", size=0.5),
        legend.text = element_text(size=5), 
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=5, vjust=0.5),
        axis.text.y=element_text(size=12),
        legend.title = element_text(size=5),
        legend.key.size = unit(0.3, 'cm'),
        legend.position = "right") +
  guides(fill=guide_legend(ncol=1))+ labs(x="", y="% taxa")+
  scale_x_discrete()

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Shared_phylum_by_site_",variation,".png"), height = 7, width=15, units="in")

```


#Get top taxa by environment  #### needs work, not finished #####
```{r}
data.ps <-ps_tr %>%
  tax_glom("Phylum") %>%
  psmelt() %>%
  as_tibble()

# highest abundance: all samples within each environment pooled together
data.ps %>%
  group_by(Phylum) %>%
  summarise(Abundance = mean(Abundance)) %>%
  group_by(sample_type) %>%
  arrange(-Abundance)%>%
  print(n=15)

```




# Notes for meeting June 27, 2024:
- May be worth removing samples that we know won't be in the analysis and repeat 1_RawSequenceProcessing step:
        - MAWC possblank samples, Calder not top and bottom sediments, anything else?
- Need to select the correct samples and rerun the ZOTUexact 2_DataCleaning step
- May be worth looking at the ZOTU99, OTU99, OTU97 for 2_DataCleaning


# Notes for meeting next, July 2024????
Do we want to merge samples by depth?
What other posibilities are there for merging samples?


# 27 June 2024 - Jordan Updating NMDS
# July 9&10, 2024 - Ioana Updating NMDS
# July 12, 2024 - Ioana Updating NMDS

