---
title: "Mountain Microbe Environment Analysis"
author: "Ioana Stefanescu"
date: "2024-07-08"
output: pdf_document
---

#1. Read in metadata and packages
```{r}
library(tidyverse)
library(ggplot2)

setwd("/Users/ioana/Desktop/MountainMicrobes/6_MetadataAnalysis")

metadata<-read_csv("Metadata.for.kept.samples.csv")
metadata<- metadata%>% filter(surface_or_deep  %in% c("surface", "deep"))

metadata$measured_temperature=as.numeric(metadata$measured_temperature)
metadata$depth_end=as.numeric(metadata$depth_end)

#get column names
#colnames(metadata)

env.vars <- metadata %>% 
        select( sample_id,sample_type,surface_or_deep,sample_type_depth,latitude,longitude,elevation,depth_start, depth_end, measured_temperature,Tann,Tmin,Tmax,MAP,sample_number) %>% 
        filter(str_detect(sample_type, "blank", negate = TRUE))  %>% 
        filter(str_detect(sample_type, "MC_71", negate = TRUE))  

```


```{r}
  custom_theme <- function() {
  theme_bw() +
    theme(
      text = element_text(color = "black", size = 13),
      axis.text = element_text(color = "black",size = 13),
      axis.title = element_text(color = "black",size = 13),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12),
      legend.text = element_text(size = 13),
      legend.title = element_text(size = 13),
      strip.text = element_text(size = 13, color = "black"),
     strip.background = element_blank(), 
                strip.text.x = element_blank())}


theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text = element_text(color = "black",size = 13),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
               ) +
```

# Correlations
```{r}
summary(lm(metadata$Tann[metadata$sample_type=="peat"]~ metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$Tann[metadata$sample_type=="soil"] ~ metadata$measured_temperature[metadata$sample_type=="soil"])) # *0.4127901 
summary(lm(metadata$Tann[metadata$sample_type=="sediment"]~ metadata$measured_temperature[metadata$sample_type=="sediment"]) )# -0.1748956 

summary(lm(metadata$Tmax[metadata$sample_type=="peat"] ~ metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$Tmax[metadata$sample_type=="soil"]~ metadata$measured_temperature[metadata$sample_type=="soil"])) # * 0.6115696
summary(lm(metadata$Tmax[metadata$sample_type=="sediment"]~ metadata$measured_temperature[metadata$sample_type=="sediment"])) # -0.171343

summary(lm(metadata$Tmin[metadata$sample_type=="peat"]~ metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$Tmin[metadata$sample_type=="soil"]~ metadata$measured_temperature[metadata$sample_type=="soil"]) )
summary(lm(metadata$Tmin[metadata$sample_type=="sediment"]~ metadata$measured_temperature[metadata$sample_type=="sediment"])) # -0.1725901 

summary(lm(metadata$MAP[metadata$sample_type=="peat"]~ metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$MAP[metadata$sample_type=="soil"]~ metadata$measured_temperature[metadata$sample_type=="soil"]) )
summary(lm(metadata$MAP[metadata$sample_type=="sediment"]~ metadata$measured_temperature[metadata$sample_type=="sediment"])) # 0.3070303 

summary(lm(metadata$elevation[metadata$sample_type=="peat"]~ metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$elevation[metadata$sample_type=="soil"]~ metadata$measured_temperature[metadata$sample_type=="soil"]) )# -0.3372595
summary(lm(metadata$elevation[metadata$sample_type=="sediment"]~ metadata$measured_temperature[metadata$sample_type=="sediment"]))


summary(lm(metadata$latitude[metadata$sample_type=="peat"]~  metadata$measured_temperature[metadata$sample_type=="peat"]))
summary(lm(metadata$latitude[metadata$sample_type=="soil"]~  metadata$measured_temperature[metadata$sample_type=="soil"]) )# -0.3057934
summary(lm(metadata$latitude[metadata$sample_type=="sediment"]~  metadata$measured_temperature[metadata$sample_type=="sediment"]))


p.t.ann=ggplot(metadata,aes(x =Tann , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "dodgerblue", "#ef767a"))+
                                         custom_theme() + labs(y="")


p.t.max=ggplot(metadata,aes(x =Tmax , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "dodgerblue", "#ef767a"))+
                                         custom_theme() + labs(y="")
 


p.t.min= ggplot(metadata,aes(x =Tmin , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "dodgerblue", "transparent"))+
                                         custom_theme() + labs(y="")
 
p.MAP=ggplot(metadata,aes(x =MAP , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "dodgerblue", "transparent"))+
                                                 custom_theme() + labs(y="")


p.elevation=ggplot(metadata,aes(x =elevation , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "transparent", "#ef767a"))+
                                         custom_theme() + labs(y="")
 

p.latitude=ggplot(metadata,aes(x =latitude , y = measured_temperature, fill = sample_type, shape = sample_type)) +
   geom_point(color="transparent", size=2) + 
        scale_shape_manual(values = c(21,22, 23))+
              scale_fill_manual(values= c("orange", "dodgerblue", "#ef767a"))+
                        geom_smooth(aes(color=sample_type),method = "lm", fill = NA, show.legend = F) +
                                scale_color_manual(values= c("transparent", "transparent", "#ef767a"))+
                                                 custom_theme() + labs(y="")
library(ggpubr)
figure1=ggarrange(p.t.ann,  p.MAP,   p.t.max, p.elevation, p.t.min,   p.latitude,
          ncol = 2, nrow = 3, 
        common.legend = TRUE, legend = "top")+ bgcolor("white")       
annotate_figure(figure1,
                left = text_grob("Substrate measured temperature", color = "black",size=25, rot = 90))

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Metadata_Correlations.png"), height = 10, width=10, units="in")



```


# Plot depths 
```{r}
metadata %>% 
        filter( sample_type == "peat") %>% 
        filter( surface_or_deep == "surface")  %>% 
        arrange(depth_start) %>%    # order based on depth start 
        ggplot()+
        geom_segment( aes(x=reorder(sample_number,depth_start), y=depth_start, yend=depth_end),size=1,  colour="#fbd1a2") +
        theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
xlab("")+ ylab("Surface samples interval (cm)")


metadata %>% 
        filter( sample_type == "sediment") %>% 
        filter( surface_or_deep == "surface")  %>% 
        arrange(depth_start) %>%    # order based on depth start 
        ggplot()+
        geom_segment( aes(x=reorder(sample_number,depth_start), y=depth_start, yend=depth_end),size=1,  colour="dodgerblue") +
        theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
xlab("")+ ylab("Surface samples interval (cm)")

```

#a. Get sample number and variable ranges
```{r}

# How many samples per 3 substrate?
metadata %>% 
    count(sample_type)

#How many samples per 5 substrates?
metadata %>% 
    count(sample_type_depth)

# How many sites?
metadata %>% 
        count(sample_type,site_name) %>%
        count(sample_type)

```



#2. Plot sample depth intervals

```{r}
env.vars %>%
        filter(str_detect(sample_type, "peat", negate = F)) %>%
        arrange(depth_start) %>%    # order based on depth start 
ggplot()+
geom_segment( aes(x=reorder(sample_number,depth_start), y=depth_start, yend=depth_end),size=1,  colour="#fbd1a2") +
theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
xlab("")+ ylab("Depth intercal (cm)")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Samples_depths_Peat.png"), height = 5, width=9, units="in")


env.vars %>%
        filter(str_detect(sample_type, "soil", negate = F)) %>%
        arrange(depth_start) %>%     # order based on depth start 
ggplot()+
geom_segment( aes(x=reorder(sample_number,depth_start), y=depth_start, yend=depth_end),size=1,  colour="#ef767a") +
theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
xlab("")+ ylab("Depth intercal (cm)")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Samples_depths_Soil.png"), height = 5, width=9, units="in")


env.vars %>%
        filter(str_detect(sample_type, "sediment", negate = F)) %>%
        arrange(depth_start) %>%     # order based on depth start 
ggplot()+
geom_segment( aes(x=reorder(sample_number,depth_start), y=depth_start, yend=depth_end),size=1,  colour="#A1C7E0") +
theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
xlab("")+ ylab("Depth intercal (cm)")

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Samples_depths_Sediment.png"), height = 5, width=9, units="in")



env.vars %>%
        mutate(numbered.rows=1:nrow(env.vars)) %>%
ggplot()+
geom_segment( aes(x=numbered.rows, y=depth_start, yend=depth_end, color=surface_or_deep),size=1) +
theme(  panel.background = element_blank(),
                text = element_text(color = "black", size = 13),
                axis.line = element_line(color="black"),
                axis.text.x  = element_blank(),
                axis.title = element_text(color = "black",size = 13),
                axis.title.x = element_text(margin = margin(t = 10)),
                axis.title.y = element_text(margin = margin(r = 10)),
                plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                plot.subtitle = element_text(hjust = 0.5, size = 12),
                legend.text = element_text(size = 10),
                legend.title = element_text(size = 13),
                strip.background = element_blank(), strip.text.x = element_blank()) +
                scale_colour_manual(values = c("grey20", "grey70")) +
        scale_y_continuous(breaks=seq(0,100, by=5))+
        labs(y="Depth interval (cm)", x="", color="")+
                geom_vline(xintercept = 271)+
                  geom_vline(xintercept = 464)+
        annotate("text", x=c(135,367, 502), y=93, label=c("Peat", "Sediment", "Soil"),  color=c("orange","dodgerblue","#ef767a"), size=4)


ggsave(device = "png",filename=paste0("../5_Figures/FigX_Samples_depths_all_types.png"), height = 5, width=9, units="in")

```

#3. Get elevation raster and US states boundries
```{r}
library(elevatr)
library(terra)

lat.min= min(env.vars$latitude, na.rm=T)
lat.max=max(env.vars$latitude, na.rm=T)
long.min= min(env.vars$longitude, na.rm=T)
long.max= max(env.vars$longitude, na.rm=T)

set.seed(65.7)
examp_df <- data.frame(x = runif(10, -111, max = -103.3), y = runif(10, 36.8,46.5))

#prj_dd <- "EPSG:4326"
prj_dd="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
df_elev_epqs <- get_elev_point(examp_df, prj = prj_dd, src = "epqs")
elevation <- get_elev_raster(df_elev_epqs, z = 9)
obj_raster <- rast(elevation) 
names(obj_raster)="elevation" # Change weird name to "elevation"

library("tidyterra")
library(sf)
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE)) # get state boundaries

```

#a. Plot sites map
```{r}
ggplot() +
  geom_spatraster(data = obj_raster, aes(fill = elevation)) +
  geom_sf(data=usa, color = "#2b2b2b", fill = "transparent", size=0.125) +
  geom_point(data=env.vars, aes(longitude, latitude,color=sample_type),shape=16,size=3)+
        scale_colour_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a")) +
        scale_fill_gradientn(colours = rev(c("grey30", "#4D4D4D",  "#838383" , "#B5B5B5" ,"grey90", "grey95")),
                limits= c(900,3850), n.breaks=10) +
        custom_theme()+
  scale_x_continuous(limits=c(-111,-104), expand=c(0,0))+
  scale_y_continuous(limits=c(37.65,45.3), expand=c(0,0))+
        labs(fill='Elevation (m)', color="Sample type") 
ggsave(device = "png",filename=paste0("../5_Figures/FigX_Samples_map.png"), height = 9, width=9, units="in")
```


#4. Plot metadata histograms: Elevation, Tmean, Tmin, Tmax, MAP
```{r}
library(tidyverse)
library(ggtext)

env.vars %>% count(sample_type)

#min(env.vars$elevation, na.rm=T)
#max(env.vars$elevation, na.rm=T)
h.elevation<- ggplot(env.vars, aes(elevation,  fill = sample_type)) + 
        geom_histogram(binwidth = (100),boundary=100, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
        custom_theme()+
scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) +       
        scale_y_continuous(expand=c(0,0))+
        xlab("Elevation (m)")+
        labs(fill='', y="") 


h.latitude<- ggplot(env.vars, aes(latitude,  fill = sample_type)) + 
        geom_histogram(binwidth = (0.5),boundary=0.5, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
               custom_theme()+
scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) +       
        scale_y_continuous(expand=c(0,0))+
        xlab("Latitude (degrees)")+
        labs(fill='', y="") 

h.t.measured<- ggplot(env.vars, aes(measured_temperature,  fill = sample_type)) + 
        geom_histogram(binwidth = (1),boundary=1, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
                custom_theme()+
scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) +       
        scale_y_continuous(expand=c(0,0))+
        xlab("Measured sample temperature (C)")+
        labs(fill='', y="") 


h.t.ann<- ggplot(env.vars, aes(Tann,  fill = sample_type)) + 
        geom_histogram(binwidth = 0.5, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
                custom_theme()+
scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) + 
        scale_y_continuous(expand=c(0,0))+
        xlab("Mean Annual Air Temperature (C)")+
        labs(fill='', y="") 


h.t.min<-ggplot(env.vars, aes(Tmin,  fill = sample_type)) + 
        geom_histogram(binwidth = 0.5, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
                custom_theme()+
        scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)"))  +       
        scale_y_continuous(expand=c(0,0))+
        xlab("Mean Annual Minimum Air temperature (C)")+
        labs(fill='', y="") 


h.t.max<-ggplot(env.vars, aes(Tmax,  fill = sample_type)) + 
        geom_histogram(binwidth = 0.5, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
                custom_theme()+
scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) +       
        scale_y_continuous(expand=c(0,0))+
        xlab("Mean Annual Maximum Air temperature (C)")+
        labs(fill='', y="") 

h.map<-ggplot(env.vars, aes(MAP,  fill = sample_type)) + 
        geom_histogram(binwidth = 50, color="black", size=0.2) +
        facet_wrap(~sample_type, ncol = 1)+
               custom_theme()+
        scale_fill_manual(values = c("#fbd1a2", "#A1C7E0", "#ef767a"),  labels=c( "peat (n=270)", "sediment (n=193)", "soil (n=77)")) +      
        scale_y_continuous(expand=c(0,0))+
        xlab("Mean Annual Precipitation (mm)")+
        labs(fill='', y="") 



library(ggpubr)
figure2 =ggarrange(h.t.ann, h.map, h.t.max,  h.elevation, h.t.min,  h.latitude, h.t.measured, 
        #  bp + rremove("x.text"), 
          ncol = 2, nrow = 4, 
        common.legend = TRUE, legend = "right")+ bgcolor("white")       
annotate_figure(figure2,
                left = text_grob("count", color = "black",size=25, rot = 90))

ggsave(device = "png",filename=paste0("../5_Figures/FigX_Metadata_histograms.png"), height = 15, width=10, units="in")

```



#Questions for next meeting
1. Should we divide the sampling depths into surface vs deep samples? How are we going to treat the different depths?

2. Dulcinea: These samples are in the 16S data but not in the Metadata: "MAWC.CR1A"        "MAWC.CR1B"        "MAWC.FR1A"        "MAWC.FR1B"        "MAWC.RR1A"        "MAWC.RR1B"  
Any toughts on why this is?

3. Jordan: These samples are in the metadata but not in the ESC file? Were all these samples removed by cleaning?
print(diffs)
 [1] "Calder.Blank3"              "Calder.JC_Blank1"           "Calder.JC_Blank2"           "Calder.JC_Blank3"          
 [5] "Calder.JC7_Blank1"          "Calder.JC7_Blank2"          "Calder.JC7_Blank3"          "Calder.JC7_Blank4"         
 [9] "Calder.JC7_Blank5"          "Calder.JC7_Blank6"          "Calder.JC7_Blank7"          "MAWC.BLANK"                
[13] "MAWC.BLANK1_DG1"            "MAWC.BLANK2_DG2"            "MAWC.BLANK3_DG3"            "MAWC.BLANK4_DG4_BLANK4_DG4"
[17] "SNOTEL.IOANA_D12"           "SNOTEL.IOANA_E12"           "SNOTEL.IOANA_F12"           "SNOTEL.IOANA_H12"          
[21] "MAWC.CC1A"                  "MAWC.LBFL21_5_14_15_cm"     "MAWC.LBFL21_5_16_17_cm"     "MAWC.LBFL21_5_18_19_cm"    
[25] "MAWC.LBFL21_5_20_21_cm"     "MAWC.LBFL21_5_22_23_cm"     "MAWC.LBFL21_5_8_9_cm"       "MAWC.LL3B_J"               
[29] "MAWC.LZ2B"                  "MAWC.ME1A_S"                "SNOTEL.IOANA_A11"           "SNOTEL.IOANA_B11"          
[33] "SNOTEL.IOANA_C10"           "SNOTEL.IOANA_C11"           "SNOTEL.IOANA_C12"           "SNOTEL.IOANA_D10"          
[37] "SNOTEL.IOANA_D11"           "SNOTEL.IOANA_E10"           "SNOTEL.IOANA_E11"           "SNOTEL.IOANA_F10"          
[41] "SNOTEL.IOANA_F11"           "SNOTEL.IOANA_G10"           "SNOTEL.IOANA_G11"           "SNOTEL.IOANA_H11"     






        
        
        

